---
title: "assign4"
author: "Shaoyi Zhang"
date: "May 24, 2016"
output: pdf_document
---
```{r global_options, include=FALSE,warning=F}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',warning=FALSE)
```


## Question 1.1
```{r}
setwd("/Users/Shawn/Desktop/PSTAT 231/PSTAT-231/assign4")
set.seed(2)
library(data.table)
food.data = read.table("food.txt",header = T,row.names = 1)
st.food = scale(food.data)
```

If we choose to have 2 clusters:

```{r}
km.2 = kmeans(food.data[2:ncol(food.data)], centers = 2, nstart = 50)

# the centroids
km.2$centers

# The size of each center
km.2$size

# The ratio of between-SS/total-SS
km.2$betweenss/km.2$totss
```

```{r}
centers = as.data.table(km.2$centers)
centers[,Group:=as.factor(c(1,2))]
library(GGally)
gpd <- ggparcoord(data = centers, columns = 1:4, groupColumn = 5)
gpd
```

Group 1 has high Protein, Fat and Iron, but Low Calcium.

Group 2 has High Calcium, but low Protein, Fat and Iron.

```{r}
# Group membership
km.2$cluster
```

From the cluster membership, we noticed that Group 2 are all Canned fish. Since seafood tends to have high Calcium, our grouping is reasonable. But there are still many "fish" categorized in Group 1 which are mostly beef/lamb/chicken.


Then, if we choose to have 3 clusters:

```{r}
km.3 = kmeans(food.data[2:ncol(food.data)], centers = 3, iter.max = 10, nstart = 50)

# the centroids
km.3$centers

# The size of each center
km.3$size

# The ratio of between-SS/total-SS
km.3$betweenss/km.2$totss
```

```{r}
centers = as.data.table(km.3$centers)
centers[,Group:=as.factor(c(1,2,3))]
library(GGally)
gpd <- ggparcoord(data = centers, columns = 1:4, groupColumn = 5)
gpd
```

Group 1 have high Fat, medium Protein, but low Calcium and Iron.

Group 2 have high Iron, medium Calcium, but low Protein and Fat.

Group 3 have high Protein and Calicium, but low Fat and Iron.


```{r}
# Group membership
km.3$cluster
```


```{r}
km.4 = kmeans(food.data[2:ncol(food.data)], centers = 4, iter.max = 10, nstart = 50)

# the centroids
km.4$centers

# The size of each center
km.4$size

# The ratio of between-SS/total-SS
km.4$betweenss/km.2$totss
```

The ratio of between-SS/total-SS didn't inrease much when we increase K = 3 to K = 4. This means 3 clusters have differentiated our data set very well. We will continue our analysis with Optimal K = 3.

Before we continue, it's also important to see the graph of the ratio of between-SS/total-SS

```{r}
library(ggplot2)
maxK = 10
st.food=scale(food.data[c(1:5)])
# k- means clustering loop
ratiowss=vector()
for (k in 2:maxK){
  km=kmeans(st.food,k,nstart=50)
  ratiowss[k]=km$tot.withinss/km$totss
}
dt=data.table("K"=2:maxK,"ratiowss"=ratiowss[2:maxK])
ggplot(dt,aes(x=K,y=ratiowss))+geom_line(size=3)
```

As we increase the number of cluster, the ratio of between-SS/total-SS will always increase. That's expected because when we add new centroids, there must be reduction in distance from some points to the new centriod.

```{r}
# use clusters = 3 for further analysis
km.out = kmeans(st.food,centers = 3)
centers = data.table(km.out$centers)
```

Let's take a look at the centers of the clusters

```{r,echo=F}
# centers
# # transpose to make it more readable
# tCenters = t(centers)
# which.max(tCenters[1,])
# 
# maxims=vector()
# for (i in 1:nrow(tCenters)){
#   maxims[i]=which.max(tCenters[i,])
#   print(which.max(tCenters[i,]))
# }
# maxims
```


## Question 1.2

Now, we will try Hierachical Clustering.

```{r,echo=F}
# compute distance matrix
dist.food = dist(st.food)

cluster.single = hclust(dist.food,method = "single")
plot(cluster.single)

cluster.complete = hclust(dist.food,method = "complete")
plot(cluster.complete)

cluster.average = hclust(dist.food,method = "average")
plot(cluster.average)
```

In min-distance method, we get the dendrogram with the lowest height.

In average-distance method, it tends to seperate one "outlier" out of the group in each split.

In max-distance method, it to have a more "cluster-like" dendrogram compared to average-distance method.

I can individuate clusters in Hierachical clustering similar in K-means clustering. However, it's more difficult now. In K-means clustering, since there aren't any "Subgroup", the interpretation is straightforward -- simply summarize the centroids. However, in Hierachical clustering, groups, except leaves, are always subgroups of more general groups. In Hierachical clustering, we can clearly see that food with high Protein and Fat are grouped together. The h-cluster algorithm also does a good job on split seafood from the others.

## Question 2
```{r}
pr.out = prcomp(food.data[1:5], center = T, scale=T)
summary(pr.out)
# The first 4 PC explained 99.97% of the variance
library(ggbiplot)
# Use ggbiplot to create PCA graph
ggbiplot(pr.out, obs.scale = 1, var.scale = 1, 
         ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = '') +
  theme(legend.direction = 'horizontal', legend.position = 'top')
pr.out
```

By PCA, we noticed that the first 4 principle components expained more than 99% of the variance in the data. And the grouping is similar to the results we obtained using other methods.

In PC1, Energy and Fat are very important. This is similar to the Hierachical clustering, which seperates high fat/energy fooe, e.g. roast beef, from other food.

In PC2, Iron and Protein are very important. This is leads to the seperation of "Clams" from other food since clams has very low Protein, but very high Iron.

In PC3, Calcium and Iron plays a big role. This distinction seperates "Canned finshes" from other food.

In PC4, Calcium and Protein explained the rest of variability.

I prefer hierachical clustering, because K-means clustering is too general and PCA is only a measure of variance which don't provide any clustering by itself.

## Question 3
```{r}
dissmilarity = matrix(c(0,0.3,0.4,0.7,0.3,0,0.5,0.8,0.4,0.5,0,0.45,0.7,0.8,0.45,0),nrow = 4)
clust.dist = hclust(as.dist(dissmilarity),method = "complete")
plot(clust.dist)
```

If we only want 2 cluster, then observation 1 and 2 will be in the same cluster, while observation 3 and 4 will be in another cluster.

```{r,echo=F}
#sorted = food.data[order(-food.data$Calcium,-food.data$Protein),]
#sorted
#colnames(food.data)
# This seperates "Canned tuna" and "Broiled mackerel" from the "beefish" food. Tuna and mackerel were originally clusterd with beef and pork because they have low Calcium and iron. However, if the Protein
```
